---
title: 'Causal Inference Final Project: Effect of Smoking on 10-year Development of
  Coronary Heart Disease'
author: "Bianca Doone, Michael Attah, Graham Casey Gibson, Daniel Saunders, Nutcha Wattanachit"
date: "November 25, 2018"
output:
  pdf_document:
           latex_engine: xelatex
           fig_caption: yes
           keep_tex: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Exploration

```{r}
fhs = read.csv('framingham.csv', header = T)
head(fhs)
```

```{r}
summary(fhs)
```

```{r}
table(fhs$TenYearCHD)
```

```{r}
table(fhs$currentSmoker)
```

```{r}
table(fhs$cigsPerDay)
```

```{r}
dim(fhs)
```

# Causal Roadmap
## Background Story

Cardiovascular disease (CVD) is the leading cause of death and serious illness in the United States. The objective of the Framingham Heart Study was to identify the common factors or characteristics that contribute to CVD by following its development over a long period of time in a large group of participants who had not yet developed overt symptoms of CVD or suffered a heart attack or stroke.

The researchers recruited 5,209 men and women between the ages of 30 and 62 from the town of Framingham, Massachusetts, and began the first round of extensive physical examinations and lifestyle interviews that they would later analyze for common patterns related to CVD development. Over the years, careful monitoring of the Framingham Study population has led to the identification of the major CVD risk factors – high blood pressure, high blood cholesterol, smoking, obesity, diabetes, and physical inactivity – as well as a great deal of valuable information on the effects of related factors such as blood triglyceride and HDL cholesterol levels, age, gender, and psychosocial issues. Although the Framingham cohort is primarily Caucasian, the importance of the major CVD risk factors identified in this group have been shown in other studies to apply almost universally among racial and ethnic groups, even though the patterns of distribution may vary from group to group.

## Step 0: Specify the Scientific Question

What is the effect of smoking on the ten-year development of Coronary Heart Disease?

#### Target population 

The target population is white middle-class men and women aged 30 to 62 in the US. 


The sameple in this study is white middle-class men and women aged 30 to 62 (at baseline) in Framingham, Massachusetts. We are willing to generalize to the target population because it is reasonable to assume that SES and risk factors among the sample and the target population are adequately homogenous.

## Step 1: Specify a Causal Model
<!--Use a SCM to represent your knowledge (and its limits) about the
system you will study. If you have uncertainty in specifying your SCM, discuss that explicitly and explain why you made the decisions you did.
NW - below is just a place holder so we can edit later-->

- Endogenous nodes: $X = (W,Z,A,Y)$, where 

- $W$ is age, gender, BMI, education, prevalence Diabetes
- $Z$ is blood pressure (systolic and diastolic), total Cholesterol, prevalence of hypertension, 
  prevalence of stroke, heart rate
- $A$ is the number of cigarettes smoked per days
- $Y$ is the ten-year development of coronary heart disease (CHD).

- Exogenous nodes: $U = (U_{W}, U_{Z}, U_A , U_Y) \sim \mathbb{P}_U$. We make no assumptions about the distribution $\mathbb{P}_U$.

- Structural equations $F$:
$$
\begin{aligned}
W1 &\leftarrow  f_{W}(U_{W})\\
Z &\leftarrow  f_{Z}(W,A,U_{Z})\\
A  &\leftarrow f_A(W,U_A)\\
Y  &\leftarrow f_Y(W,Z,A,U_Y)\\
\end{aligned}
$$

There are no exclusion restrictions or assumptions about functional form.


## Step 2: Counterfactuals & Causal Parameter

#### Causal Parameter
$${\Psi^*}^{i}(\mathbb{P}^*)=\mathbb{E}^*[Y_i]\ \ \ i\in \{1,2,3,4}$$
where $i$ represent the bin of cigarettes smoked per day. $Y_i$ denotes the counterfactual outcome (the ten-year development of cardiovascular disease), if possibly contrary to fact, a person's number of cigarettes smoked per day is within $i^{th}$ bin. 

<!--#### G-Computation
Here we let $\mathbb{W} = \{W_1,W_2\}$
$$\Psi_O(\mathbb{P}^i_O)=\mathbb{E}_o[\mathbb{E}_o[Y|A=a \text{ in bin } i,\mathbb{W}]]$$

$$\Psi_n(\mathbb{P}^i_n) = \frac{1}{n}\sum_{j=1}^n \mathbb{E}_n(Y|A=a \text{ in bin } i,\mathbb{W})$$
where $j$ indexes 1 to (TBD) observations. 

#### IPTW

$$IPTW^i = \frac{1}{n} \sum_{j}^n Y\frac{ \mathbb{I} (A \in i)}{P(A \in i | \mathbb{W})}$$ 

#### Targeted Maximum Likelihood Estimate (TMLE)-->

## Step 3. Specify your observed data and its link to the causal model

- Describe your observed data and its link to the causal model you have specified. If you feel that in reality the link between your causal model and the observed data is more complex than we have learned in class (n i.i.d. copies of random variable O), explain why. But for this project, stick with the simple link we have learned in class.


The dataset is adapted from Framingham Heart Study. We assume that Gender, Age, Education, BMI, Diabetes Status, and Number of Cigarettes per Day ($A$) were collected in a questionnaire at baseline. Then, Prevalence of Stroke, Prevalence of Hypertension, Indication of Blood Pressure Medication, Total Cholesterol Level, Blood Pressure, and Heart Rate were all collected after the questionnaire at a doctor's office. Our outcome, Coronary Heart Disease, is collected at a 10-year follow up. Note that this is unlike the original study. We assume our observed data were generated by sampling $n$ from a system described by our structural causal model, so we have $n = 4211$ copies of $O\overset{i.i.d}\sim \mathbb{P}_O$. We place no restrictions on the statistical model $\mathcal{M}$, which is thereby non-parametric.

\begin{table}[ht]
\centering
\caption{Descriptive Table}
\begin{tabular}{|l|l|l|}
\hline
Variable Name & Covariate & Description \\ \hline
male & Gender & binary: \ male = 1 \ female = 0 \\ \hline
age & Age & ordinal: \ 32-38, 39-40, 41-43, 44-45, 46-48 \\
& & 49-51, 52-54, 55-57, 58-61, 62-70 \\ \hline
education & Education level & ordinal: \ 1 = some high school, 2 = high school/GED \\
& & 3 = some college/vocational school, 4 = college \\ \hline
currentSmoker & Current Smoking Status & binary: \ 1 = Yes \ 0 = No \\ \hline
cigsPerDay & Number of cigarettes per day & ordinal: \ 0, 1-19, 20-70 \\ \hline
BPMeds & Indicator of blood pressure medication & binary: \ 1 = Yes \  0 = No \\ \hline
prevalentStroke & Prevalence of Stroke & binary: \ 1 = Yes \ 0 = No \\ \hline
prevalentHyp & Prevalence of Hypertension & binary: \ 1 = Yes \ 0 = No \\ \hline
diabetes & Prevalence of Diabetes & binary: \ 1 = Yes \ 0 = No \\ \hline
totChol & Total Cholesterol Level & ordinal: \ 0-79, 80-89, 90-600 \\ \hline
sysBP & Systolic Blood Pressure & ordinal: \ 0-119, 120-139, 140-295 \\ \hline
diaBP & Diastolic Blood Pressure & ordinal: \ 0-79, 80-89, 90-142.5  \\ \hline
BMI & Body Mass Index & ordinal: \ 0-18.4, 18.5-24.9, 25-29.9, 30-56.8 \\ \hline
heartRate & Heart Rate & ordinal: \ 0-59, 60 -143 \\ \hline
glucose & Glucose Level & ordinal: \ 0-77, 78-394 \\ \hline
TenYearCHD & Ten Year Follow-Up Prevalence & binary: \ 1 = Yes (had CHD) \\
& & 0 = No (do not have CHD) \\ \hline
\end{tabular}
\end{table}


```{r, include = FALSE}
library(dplyr)

### REMOVE EXPOSURE AND OUTCOME NA
fhs <- fhs %>%  filter(!is.na(cigsPerDay),!is.na(TenYearCHD))

```


 
```{r, include = FALSE}
colnames(fhs)
library(mltools)

## initalize empty binned dataframe 
fhs_binned <- data.frame(i=1:nrow(fhs))

## cigsPerDay
fhs_binned$cigsPerDay <-bin_data(fhs$cigsPerDay,bins=c(0,1,10,20,max(fhs$cigsPerDay)), binType = "explicit")

## diabetes
fhs_binned$diabetes <- as.factor(fhs$diabetes)

## Stroke
fhs_binned$prevalentStroke <- as.factor(fhs$prevalentStroke)

## Hypertension
fhs_binned$prevalentHyp <- as.factor(fhs$prevalentHyp)


# Variables binned based on science!
## age
fhs_binned$age  <- bin_data(fhs$age, bins=4, binType = "quantile")

## education NA fix
fhs[is.na(fhs$education),]$education <- 9
fhs_binned$education  <- (fhs$education)

## making 1 BP var
fhs_binned$BP <- as.factor(ifelse(fhs$sysBP/fhs$diaBP < 128/80,0,1))

## total cholesterol
fhs_binned$totChol  <- bin_data(fhs$totChol, bins=c(0,200,240,max(fhs$totChol)), binType = "explicit")

## gender
fhs_binned$gender  <-factor(fhs$male)

## BMI NA fix
fhs[is.na(fhs$BMI),]$BMI <- 9
fhs_binned$bmi  <- bin_data(fhs$BMI, bins=c(0,18.5,25,30,max(fhs$BMI)), binType = "explicit")

## technically don't need this...
fhs_binned$glucose  <- bin_data(fhs$glucose, bins=c(0,78,max(fhs$glucose)), binType = "explicit")

## heart rate
fhs_binned$heartRate  <- bin_data(fhs$heartRate, bins=c(0,60,max(fhs$heartRate)), binType = "explicit")

## CHD
fhs_binned$CHD <- factor(fhs$TenYearCHD)

## remove index created 
fhs_binned <- subset(fhs_binned, select = -c(i))
```

## Step 4. Identifiability
Is your target causal parameter identified under your initial causal model? If not, under what additional assumptions would it be identified? How plausible are these for your particular problem? Are there additional data or changes to your study design that would improve their plausibility?

## Step 5. Statistical Model and Estimand

The target parameter of $\mathbb{P}_0$, which equals the causal parameter in the augmented causal model $\mathcal{M}^{**}$ is given by the G-Computation formula:


$$
\begin{aligned}
\Psi_0(\mathbb{P}^i_0)&=\mathbb{E}_0[\mathbb{E}_0[Y|A= i,W=w]]\\
&= \sum_w\mathbb{E}_0[Y|A=i,W=w]*\mathbb{P}_0(W=w)
\end{aligned}
$$

<!--$$\Psi_n(\mathbb{P}^i_n) = \frac{1}{n}\sum_{j=1}^n \mathbb{E}_n(Y|A=a \text{ in bin } i,\mathbb{W})$$
where $j$ indexes 1 to (TBD) observations.--> 

## Step 6. Estimation

### Conditional Mean outcome 

```{r}
library(mgcv)
glm_fit <- glm( CHD ~ cigsPerDay + education + age + diabetes + bmi , data = fhs_binned, family = "binomial")
summary(glm_fit)

intervene_on_bin <- function(i){
  fhs_binned_i <- fhs_binned
  fhs_binned_i$cigsPerDay <-levels(fhs_binned$cigsPerDay)[i]
  return (fhs_binned_i)
}


average_treatment_effect <- c()
average_treatment_effect_ci <- matrix(NA,nrow=length(levels(fhs_binned$cigsPerDay)),ncol=2)
for (i in 1:length(levels(fhs_binned$cigsPerDay))){
  average_treatment_effect[i] <- mean(predict(glm_fit, newdata=intervene_on_bin(i), type='response'))
  average_treatment_effect_ci[i,] <- quantile(predict(glm_fit, newdata=intervene_on_bin(i), type='response'),probs=c(.025,.975))

}

plot(average_treatment_effect,ylab="Probability of CHD",ylim=c(0,.5))
points(average_treatment_effect_ci[,1],col='red',lty=2)
points(average_treatment_effect_ci[,2],col='red',lty=2)
```

### IPTW
```{r}

### Create pairwise binary variables for each bin
fhs_binned$cigsPerDay_bin_1 <- ifelse(fhs_binned$cigsPerDay == "[0, 1)",1,0)
fhs_binned$cigsPerDay_bin_2 <- ifelse(fhs_binned$cigsPerDay == "[1, 10)",1,0)
fhs_binned$cigsPerDay_bin_3 <- ifelse(fhs_binned$cigsPerDay == "[10, 20)",1,0)
fhs_binned$cigsPerDay_bin_4 <- ifelse(fhs_binned$cigsPerDay == "[20, 70]",1,0)

### BIN 1
glm_fit_iptw_bin_1 <- glm( cigsPerDay_bin_1 ~   education + age + diabetes + bmi , data = fhs_binned, family = "binomial")
prob.1W <- predict(glm_fit_iptw_bin_1, type= "response")
wt_1<- 1/prob.1W
summary(wt_1)
IPTW_bin_1<- mean( wt_1*as.numeric(fhs_binned$cigsPerDay_bin_1==1)*as.numeric(fhs_binned$CHD==1))
IPTW_bin_1_ci <- quantile(wt_1*as.numeric(fhs_binned$cigsPerDay_bin_1==1)*as.numeric(fhs_binned$CHD==1),probs=c(.025,.975))

### BIN 2
glm_fit_iptw_bin_2 <- glm( cigsPerDay_bin_2 ~   education + age + diabetes + bmi , data = fhs_binned, family = "binomial")
prob.1W <- predict(glm_fit_iptw_bin_2, type= "response")
wt_2<- 1/prob.1W
summary(wt_2)
IPTW_bin_2<- mean( wt_2*as.numeric(fhs_binned$cigsPerDay_bin_2==1)*as.numeric(fhs_binned$CHD==1))
IPTW_bin_2_ci <- quantile(wt_2*as.numeric(fhs_binned$cigsPerDay_bin_2==1)*as.numeric(fhs_binned$CHD==1),probs=c(.025,.975))

### BIN 3
glm_fit_iptw_bin_3 <- glm( cigsPerDay_bin_3 ~   education + age + diabetes + bmi , data = fhs_binned, family = "binomial")
prob.1W <- predict(glm_fit_iptw_bin_3, type= "response")
wt_3<- 1/prob.1W
summary(wt_3)
IPTW_bin_3<- mean( wt_3*as.numeric(fhs_binned$cigsPerDay_bin_3==1)*as.numeric(fhs_binned$CHD==1))
IPTW_bin_3_ci <- quantile((wt_3*as.numeric(fhs_binned$cigsPerDay_bin_3==1)*as.numeric(fhs_binned$CHD==1)),probs=c(.025,.975))

### BIN 4
glm_fit_iptw_bin_4 <- glm( cigsPerDay_bin_4 ~   education + age + diabetes + bmi , data = fhs_binned, family = "binomial")
prob.1W <- predict(glm_fit_iptw_bin_4, type= "response")
wt_4<- 1/prob.1W
summary(wt_4)
IPTW_bin_4<- mean( wt_4*as.numeric(fhs_binned$cigsPerDay_bin_4==1)*as.numeric(fhs_binned$CHD==1))
IPTW_bin_4_ci <- quantile(wt_4*as.numeric(fhs_binned$cigsPerDay_bin_4==1)*as.numeric(fhs_binned$CHD==1),probs=c(.025,.975))

plot(c(IPTW_bin_1,IPTW_bin_2,IPTW_bin_3,IPTW_bin_4),type='l',col='blue',ylab="IPTW Estimate of CHD",xlab="Bin")
lines(IPTW_bin_1_ci)
lines(c(IPTW_bin_1_ci[1],IPTW_bin_2_ci[1],IPTW_bin_3_ci[1],IPTW_bin_4_ci[1]), type = "l", col = "red")
lines(c(IPTW_bin_1_ci[2],IPTW_bin_2_ci[2],IPTW_bin_3_ci[2],IPTW_bin_4_ci[2]), type = "l", col = "red")
```


### Superlearner/TMLE

```{r}

library('SuperLearner')
SL.library<- c("SL.glmnet","SL.randomForest","SL.nnet","SL.earth","SL.bayesglm")#,"SL.earth")

n_bootstrap_samples <-10
tmle_bootstrap <- matrix(NA, nrow=n_bootstrap_samples,ncol=4)

for (i in 1:n_bootstrap_samples){
  
  ### BIN1 TMLE
  fhs_binned_sample <- fhs_binned[sample(1:nrow(fhs_binned),nrow(fhs_binned),replace=TRUE),]
  
  X_minus_bin_1<- subset(fhs_binned_sample, select= c("cigsPerDay_bin_1", "education", "age","diabetes","bmi") )
  X_minus_bin_1$age <- as.numeric(X_minus_bin_1$age)
  X_minus_bin_1$bmi <- as.numeric(X_minus_bin_1$bmi)
  X_minus_bin_1_all_bin_1 <- X_minus_bin_1
  X_minus_bin_1_all_bin_1$cigsPerDay_bin_1 <- 1
  
  ##CONVERT FACTORS TO NUMERIC FOR SUPERLEARNER
  
  
  SL.outcome<- SuperLearner(Y=as.numeric(fhs_binned$CHD==1), X=X_minus_bin_1, SL.library=SL.library, family="binomial")
  
  
  expY.givenA1 <-  predict(SL.outcome, newdata=X_minus_bin_1_all_bin_1)$pred
  SL.exposure<- SuperLearner(Y=as.numeric(fhs_binned$cigsPerDay_bin_1==1), X=subset(X_minus_bin_1, select= -c(cigsPerDay_bin_1)),SL.library=SL.library, family="binomial")
  probA1.givenW<- SL.exposure$SL.predict
  H.AW<- as.numeric(fhs_binned$cigsPerDay_bin_1==1)/probA1.givenW
  logitUpdate<- glm(fhs_binned$CHD ~ -1 +offset(qlogis(expY.givenA1)) + H.AW, family='binomial')
  epsilon<- logitUpdate$coef
  expY.givenAW.star<- plogis(qlogis(expY.givenA1)+ epsilon*H.AW)
  PsiHat.TMLE_bin_1<- mean(expY.givenAW.star)#- expY.given0W.star)
  
  
  #### BIN 2 TMLE
  
  X_minus_bin_2<- subset(fhs_binned_sample, select= c("cigsPerDay_bin_2", "education", "age","diabetes","bmi") )
  X_minus_bin_2$age <- as.numeric(X_minus_bin_2$age)
  X_minus_bin_2$bmi <- as.numeric(X_minus_bin_2$bmi)
  X_minus_bin_2_all_bin_2 <- X_minus_bin_2
  X_minus_bin_2_all_bin_2$cigsPerDay_bin_2 <- 1
  
  SL.outcome<- SuperLearner(Y=as.numeric(fhs_binned$CHD==1), X=X_minus_bin_2, SL.library=SL.library, family="binomial")
  expY.givenA1 <-  predict(SL.outcome, newdata=X_minus_bin_2_all_bin_2)$pred
  SL.exposure<- SuperLearner(Y=as.numeric(fhs_binned$cigsPerDay_bin_2==1), X=subset(X_minus_bin_2, select= -c(cigsPerDay_bin_2)),SL.library=SL.library, family="binomial")
  probA1.givenW<- SL.exposure$SL.predict
  H.AW<- as.numeric(fhs_binned$cigsPerDay_bin_2==1)/probA1.givenW
  logitUpdate<- glm(fhs_binned$CHD ~ -1 +offset(qlogis(expY.givenA1)) + H.AW, family='binomial')
  epsilon<- logitUpdate$coef
  expY.givenAW.star<- plogis(qlogis(expY.givenA1)+ epsilon*H.AW)
  PsiHat.TMLE_bin_2<- mean(expY.givenAW.star)#- expY.given0W.star)
  
  
  
  #### BIN 3 TMLE 
  
  X_minus_bin_3<- subset(fhs_binned_sample, select= c("cigsPerDay_bin_3", "education", "age","diabetes","bmi") )
  X_minus_bin_3$age <- as.numeric(X_minus_bin_3$age)
  X_minus_bin_3$bmi <- as.numeric(X_minus_bin_3$bmi)
  
  X_minus_bin_3_all_bin_3 <- X_minus_bin_3
  X_minus_bin_3_all_bin_3$cigsPerDay_bin_3 <- 1
  SL.outcome<- SuperLearner(Y=as.numeric(fhs_binned$CHD==1), X=X_minus_bin_3, SL.library=SL.library, family="binomial")
  expY.givenA1 <-  predict(SL.outcome, newdata=X_minus_bin_3_all_bin_3)$pred
  SL.exposure<- SuperLearner(Y=as.numeric(fhs_binned$cigsPerDay_bin_3==1), X=subset(X_minus_bin_3, select= -c(cigsPerDay_bin_3)),SL.library=SL.library, family="binomial")
  probA1.givenW<- SL.exposure$SL.predict
  H.AW<- as.numeric(fhs_binned$cigsPerDay_bin_3==1)/probA1.givenW
  logitUpdate<- glm(fhs_binned$CHD ~ -1 +offset(qlogis(expY.givenA1)) + H.AW, family='binomial')
  epsilon<- logitUpdate$coef
  expY.givenAW.star<- plogis(qlogis(expY.givenA1)+ epsilon*H.AW)
  PsiHat.TMLE_bin_3 <- mean(expY.givenAW.star)
  
  
  #### BIN 4  TMLE 
  
  X_minus_bin_4<- subset(fhs_binned_sample, select= c("cigsPerDay_bin_4", "education", "age","diabetes","bmi") )
  X_minus_bin_4$age <- as.numeric(X_minus_bin_4$age)
  X_minus_bin_4$bmi <- as.numeric(X_minus_bin_4$bmi)
  
  X_minus_bin_4_all_bin_4 <- X_minus_bin_4
  X_minus_bin_4_all_bin_4$cigsPerDay_bin_4 <- 1
  SL.outcome<- SuperLearner(Y=as.numeric(fhs_binned$CHD==1), X=X_minus_bin_4, SL.library=SL.library, family="binomial")
  expY.givenA1 <-  predict(SL.outcome, newdata=X_minus_bin_4_all_bin_4)$pred
  SL.exposure<- SuperLearner(Y=as.numeric(fhs_binned$cigsPerDay_bin_4==1), X=subset(X_minus_bin_4, select= -c(cigsPerDay_bin_4)),SL.library=SL.library, family="binomial")
  probA1.givenW<- SL.exposure$SL.predict
  H.AW<- as.numeric(fhs_binned$cigsPerDay_bin_4==1)/probA1.givenW
  logitUpdate<- glm(fhs_binned$CHD ~ -1 +offset(qlogis(expY.givenA1)) + H.AW, family='binomial')
  epsilon<- logitUpdate$coef
  expY.givenAW.star<- plogis(qlogis(expY.givenA1)+ epsilon*H.AW)
  PsiHat.TMLE_bin_4 <- mean(expY.givenAW.star)
  
  
  tmle_bootstrap[i,] <- c(PsiHat.TMLE_bin_1,PsiHat.TMLE_bin_2,PsiHat.TMLE_bin_3,PsiHat.TMLE_bin_4)
}
```


```{r}

average_treatment_effect <- c()
average_treatment_effect_ci <- matrix(NA,nrow=length(levels(fhs_binned$cigsPerDay)),ncol=2)
for (i in 1:length(levels(fhs_binned$cigsPerDay))){
  average_treatment_effect[i] <- colMeans(tmle_bootstrap)[i]
  average_treatment_effect_ci[i,] <- quantile(tmle_bootstrap[,i],probs=c(.025,.975))

}

plot(average_treatment_effect,ylab="Probability of CHD",ylim=c(0,.5))
points(average_treatment_effect_ci[,1],col='red',lty=2)
points(average_treatment_effect_ci[,2],col='red',lty=2)

```

## Step 7. Result Interpretation

What is the statistical interpretation of your analyses? Discuss differences (or lack thereof) in the estimates provided by the different estimators. What is the causal interpretation of your results and how plausible is it? What are key limitations of your analysis? How might these results (if at all) inform policy, understanding, and/or the design of future studies?


```{r}

mean(as.numeric(fhs_binned[fhs_binned$cigsPerDay == "[10, 20)" & fhs_binned$age == "[32, 42)",]$CHD)-1)

mean(as.numeric(fhs_binned[fhs_binned$cigsPerDay == "[0, 1)" & fhs_binned$age == "[32, 42)",]$CHD)-1)

mean(as.numeric(fhs_binned[fhs_binned$cigsPerDay == "[1, 10)" & fhs_binned$age == "[56, 70]",]$CHD)-1)

mean(as.numeric(fhs_binned[fhs_binned$cigsPerDay == "[10, 20)" & fhs_binned$age == "[56, 70]",]$CHD)-1)
